---
title: "create_stockRelation"
author: "Kirsten Birch Håkansson, DTU Aqua"
date: "`r Sys.Date()`"
output: html_document
---

# Introduction


For now this script will always start with the relation created in fishPi2 many years ago, since a lot of work was put into creating the relation and it is believed more complete than ICES_StockCode and we already have code for improving it.

ICES ICES_StockCode has issues with incomplete list of areas. Futher, it does not include Fisheries management units (FMU) like FU for Nephrop and SAN for sandeel.

icesSD::getSD will be used for getting correct stock and species codes, plus other relevant information.

DC_Annex_1 from ICES big data call will be checked, so it is possible to join to it.

When ICES has created their stock relation it will be based on that one.

The relation will hold the following fields
stock_sid: the stock code used in ICES SID, from icesSD::getSD, this is considred the correct one
stock_annex_1: should be equal to stock_sid, but somethime
species_aphia:
species_fao:
species_scientific_name:
area: 
fmu:
rectangle:
quarter:
commment:


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F)

library(icesVocab)
library(icesSD)
library(dplyr)

update_year <- 2026

devtools::source_url("https://github.com/ices-tools-dev/RDBESstockCoord/blob/main/R/funMakeRelation.R?raw=TRUE")
```

# Before updating,  do the following

1.  Find the newest version of the DC_Annex_1.xlsx and copy it to inst\extdata

# Getting code list from ICES

```{r}
stocks_ices <- icesVocab::getCodeList("ICES_StockCode")

stocks_sid <- icesSD::getSD(year = 2026)

stocks_data_call <- c()

codes_aph <- icesVocab::getCodeList("SpecWoRMS")
  names(codes_aph)[names(codes_aph) == "Key"] <- "speciesCode"
  names(codes_aph)[names(codes_aph) == "Description"] <- "SpeciesName"

  codes_FAO <- icesVocab::getCodeList("SpecASFIS")
  names(codes_FAO)[names(codes_FAO) == "Key"] <- "Species"
  names(codes_FAO)[names(codes_FAO) == "Description"] <- "SpeciesName"

  codes_aph_FAO <- merge(codes_FAO, codes_aph, by = "SpeciesName")
  codes_aph_FAO <- subset(codes_aph_FAO, 
                          select=c("Species",
                                   "SpeciesName",
                                   "speciesCode"))

ices_area <- icesVocab::getCodeList("ICES_Area")
ices_area <- subset(ices_area, substr(Key, 1, 3) == "27." & Deprecated == F)

for (i in c(2:16)) {
  
  dat_0 <- readxl::read_xlsx(paste0("../inst/extdata/", "DC_Annex_1.xlsx"), sheet = c(i))
  
  stocks_data_call <- rbind(stocks_data_call, dat_0)
  
  rm(dat_0)
}

stocks <- funMakeRelation(year = 2026)
```

# Fix stock per WG
```{r}
unique(stocks$EG)
```

## HAWG
```{r}
wg_name <- "HAWG"
wg <- subset(stocks, EG == wg_name)
wg_dc <- subset(stocks_data_call, `Expert Group` == wg_name)
wg_eg <- subset(stocks_sid, ExpertGroup == wg_name)

# Check from stock with overlapping areas, species
dup <- mutate(group_by(wg, SpeciesName, ICESArea), no = length(StockCode))
length(subset(dup, no > 1))

subset(select(dup, -StockKeyDescription), no > 1)

## Remove duplicate no > 1
wg <- subset(wg, !(StockCode == "her.27.3a47d" & ICESArea %in% c("27.3.a", "27.3.a.20", "27.3.a.21")))

rm(dup)

## Remove other problems
wg <- subset(wg, !(StockCode == "ple.27.21-32" & ICESArea == "27.3.a.20"))

# Check that all data call and sid stocks are included
mis_eg <- subset(wg_eg, !(StockKeyLabel %in% wg$StockCode))
unique(mis_eg$StockKeyLabel)
mis_dc <- subset(wg_dc, !(`Stock Key Label` %in% wg$StockCode))
unique(mis_dc$`Stock Key Label`)

## Add missing stocks
stock <- unique(mis_eg$StockKeyLabel)[1]
san.sa.5r <- data.frame(
  SpeciesName = rep(mis_eg$SpeciesScientificName[mis_eg$StockKeyLabel == stock], 1),
  StockCode = rep(stock, 1),
  EG = rep(mis_eg$ExpertGroup[mis_eg$StockKeyLabel == stock], 1),
  StockKey = rep(mis_eg$StockKey[mis_eg$StockKeyLabel == stock], 1),
  StockKeyDescription = rep(mis_eg$StockKeyDescription[mis_eg$StockKeyLabel == stock], 1),
  ICESArea = c("27.4.a"),
  Species = rep("SAN", 1),
  speciesCode = rep("1758", 1)
)

stock <- unique(mis_eg$StockKeyLabel)[2]
san.sa.6 <- data.frame(
  SpeciesName = rep(mis_eg$SpeciesScientificName[mis_eg$StockKeyLabel == stock], 5),
  StockCode = rep(stock, 5),
  EG = rep(mis_eg$ExpertGroup[mis_eg$StockKeyLabel == stock], 5),
  StockKey = rep(mis_eg$StockKey[mis_eg$StockKeyLabel == stock], 5),
  StockKeyDescription = rep(mis_eg$StockKeyDescription[mis_eg$StockKeyLabel == stock], 5),
  ICESArea = c("27.3.a", "27.3.a.20", "27.3.a.21", "27.3.c", "27.3.c.22"),
  Species = rep("SAN", 5),
  speciesCode = rep("1758", 5)
)

stock <- unique(mis_eg$StockKeyLabel)[3]
san.sa.7r <- data.frame(
  SpeciesName = rep(mis_eg$SpeciesScientificName[mis_eg$StockKeyLabel == stock], 1),
  StockCode = rep(stock, 1),
  EG = rep(mis_eg$ExpertGroup[mis_eg$StockKeyLabel == stock], 1),
  StockKey = rep(mis_eg$StockKey[mis_eg$StockKeyLabel == stock], 1),
  StockKeyDescription = rep(mis_eg$StockKeyDescription[mis_eg$StockKeyLabel == stock], 1),
  ICESArea = c("27.4.a"),
  Species = rep("SAN", 1),
  speciesCode = rep("1758", 1)
)

stock <- unique(mis_eg$StockKeyLabel)[4]
san.sa.1r <- data.frame(
  SpeciesName = rep(mis_eg$SpeciesScientificName[mis_eg$StockKeyLabel == stock], 2),
  StockCode = rep(stock, 2),
  EG = rep(mis_eg$ExpertGroup[mis_eg$StockKeyLabel == stock], 2),
  StockKey = rep(mis_eg$StockKey[mis_eg$StockKeyLabel == stock], 2),
  StockKeyDescription = rep(mis_eg$StockKeyDescription[mis_eg$StockKeyLabel == stock], 2),
  ICESArea = c("27.4.b", "27.4.c"),
  Species = rep("SAN", 2),
  speciesCode = rep("1758", 2)
)

stock <- unique(mis_eg$StockKeyLabel)[5]
san.sa.2r <- data.frame(
  SpeciesName = rep(mis_eg$SpeciesScientificName[mis_eg$StockKeyLabel == stock], 3),
  StockCode = rep(stock, 3),
  EG = rep(mis_eg$ExpertGroup[mis_eg$StockKeyLabel == stock], 3),
  StockKey = rep(mis_eg$StockKey[mis_eg$StockKeyLabel == stock], 3),
  StockKeyDescription = rep(mis_eg$StockKeyDescription[mis_eg$StockKeyLabel == stock], 3),
  ICESArea = c("27.4.b", "27.4.c", "27.3.a.20"),
  Species = rep("SAN", 3),
  speciesCode = rep("1758", 3)
)

stock <- unique(mis_eg$StockKeyLabel)[6]
san.sa.3r <- data.frame(
  SpeciesName = rep(mis_eg$SpeciesScientificName[mis_eg$StockKeyLabel == stock], 3),
  StockCode = rep(stock, 3),
  EG = rep(mis_eg$ExpertGroup[mis_eg$StockKeyLabel == stock], 3),
  StockKey = rep(mis_eg$StockKey[mis_eg$StockKeyLabel == stock], 3),
  StockKeyDescription = rep(mis_eg$StockKeyDescription[mis_eg$StockKeyLabel == stock], 3),
  ICESArea = c("27.4.b", "27.4.c", "27.3.a.20"),
  Species = rep("SAN", 3),
  speciesCode = rep("1758", 3)
)

stock <- unique(mis_eg$StockKeyLabel)[7]
san.sa.4 <- data.frame(
  SpeciesName = rep(mis_eg$SpeciesScientificName[mis_eg$StockKeyLabel == stock], 2),
  StockCode = rep(stock, 2),
  EG = rep(mis_eg$ExpertGroup[mis_eg$StockKeyLabel == stock], 2),
  StockKey = rep(mis_eg$StockKey[mis_eg$StockKeyLabel == stock], 2),
  StockKeyDescription = rep(mis_eg$StockKeyDescription[mis_eg$StockKeyLabel == stock], 2),
  ICESArea = c("27.4.b", "27.4.c"),
  Species = rep("SAN", 2),
  speciesCode = rep("1758", 2)
)

stock <- unique(mis_eg$StockKeyLabel)[8]
san.27.6a <- data.frame(
  SpeciesName = rep(mis_eg$SpeciesScientificName[mis_eg$StockKeyLabel == stock], 1),
  StockCode = rep(stock, 1),
  EG = rep(mis_eg$ExpertGroup[mis_eg$StockKeyLabel == stock], 1),
  StockKey = rep(mis_eg$StockKey[mis_eg$StockKeyLabel == stock], 1),
  StockKeyDescription = rep(mis_eg$StockKeyDescription[mis_eg$StockKeyLabel == stock], 1),
  ICESArea = c("27.6.a"),
  Species = rep("SAN", 1),
  speciesCode = rep("1758", 1)
)

rm(mis_eg, mis_dc)

wg <- rbind(wg, san.27.6a, san.sa.1r, san.sa.2r, san.sa.3r, san.sa.4,
            san.sa.5r, san.sa.6, san.sa.7r)
rm(san.27.6a, san.sa.1r, san.sa.2r, san.sa.3r, san.sa.4,
            san.sa.5r, san.sa.6, san.sa.7r)

hawg_ok <- distinct(wg, SpeciesName, StockCode, EG, StockKey, 
StockKeyDescription, ICESArea, Species, speciesCode)

# Check correct areas
hawg_tab <- dplyr::distinct(hawg_ok, SpeciesName, StockCode, ICESArea, StockKey)

hawg_tab_t <- tidyr::spread(hawg_tab, key = ICESArea, value = StockKey)

rm(hawg_tab, wg_dc, wg_eg, wg, hawg_tab_t)
```


## WGBFAS
```{r}
wg_name <- "WGBFAS"
wg <- subset(stocks, EG == wg_name)
wg_dc <- subset(stocks_data_call, `Expert Group` == wg_name)
wg_eg <- subset(stocks_sid, ExpertGroup == wg_name)

# Check from stock with overlapping areas, species
dup <- mutate(group_by(wg, SpeciesName, ICESArea), no = length(StockCode))
length(subset(dup, no > 1))

subset(select(dup, -StockKeyDescription), no > 1)

## Remove duplicate no > 1
wg <- subset(wg, !(StockCode == "cod.27.24-32" & ICESArea == "27.3.d.24"))
wg <- subset(wg, !(Species == "QDQ"))

rm(dup)

## Remove other problems
wg <- subset(wg, !(StockCode == "ple.27.21-32" & ICESArea == "27.3.a.20"))

# Check that all data call and sid stocks are included
mis_eg <- subset(wg_eg, !(StockKeyLabel %in% wg$StockCode))
unique(mis_eg$StockKeyLabel)
mis_dc <- subset(wg_dc, !(`Stock Key Label` %in% wg$StockCode))
unique(mis_dc$`Stock Key Label`)

## Add missing stocks
stock <- unique(mis_dc$`Stock Key Label`)[1]
bzq.27.2425 <- data.frame(
  SpeciesName = rep(mis_eg$SpeciesScientificName[mis_eg$StockKeyLabel == stock], 2),
  StockCode = rep(stock, 2),
  EG = rep(mis_eg$ExpertGroup[mis_eg$StockKeyLabel == stock], 2),
  StockKey = rep(mis_eg$StockKey[mis_eg$StockKeyLabel == stock], 2),
  StockKeyDescription = rep(mis_eg$StockKeyDescription[mis_eg$StockKeyLabel == stock], 2),
  ICESArea = c("27.3.d.24", "27.3.d.25"),
  Species = rep("BZQ", 2),
  speciesCode = rep("126119", 2)
)

bzq.27.2425$SpeciesName <- paste(bzq.27.2425$SpeciesName, "Platichthys flesus", "Platichthys solemdali", sep = ",")
bzq.27.2425 <- tidyr::separate_longer_delim(data = bzq.27.2425, cols = SpeciesName, ",")

bzq.27.2425$speciesCode[bzq.27.2425$SpeciesName == "Platichthys flesus"] <- "127141"
bzq.27.2425$speciesCode[bzq.27.2425$SpeciesName == "Platichthys solemdali"] <- "1264298"

stock <- unique(mis_dc$`Stock Key Label`)[2]
bzq.27.2628 <- data.frame(
  SpeciesName = rep(mis_eg$SpeciesScientificName[mis_eg$StockKeyLabel == stock], 4),
  StockCode = rep(stock, 4),
  EG = rep(mis_eg$ExpertGroup[mis_eg$StockKeyLabel == stock], 4),
  StockKey = rep(mis_eg$StockKey[mis_eg$StockKeyLabel == stock], 4),
  StockKeyDescription = rep(mis_eg$StockKeyDescription[mis_eg$StockKeyLabel == stock], 4),
  ICESArea = c("27.3.d.26", "27.3.d.28", "27.3.d.28.1", "27.3.d.28.2"),
  Species = rep("BZQ", 4),
  speciesCode = rep("126119", 4)
)

bzq.27.2628$SpeciesName <- paste(bzq.27.2628$SpeciesName, "Platichthys flesus", "Platichthys solemdali", sep = ",")
bzq.27.2628 <- tidyr::separate_longer_delim(data = bzq.27.2628, cols = SpeciesName, ",")

bzq.27.2628$speciesCode[bzq.27.2628$SpeciesName == "Platichthys flesus"] <- "127141"
bzq.27.2628$speciesCode[bzq.27.2628$SpeciesName == "Platichthys solemdali"] <- "1264298"

rm(mis_eg, mis_dc)

wg <- rbind(wg, bzq.27.2425, bzq.27.2628)
rm(bzq.27.2425, bzq.27.2628)

wgbfas_ok <- distinct(wg, SpeciesName, StockCode, EG, StockKey, 
StockKeyDescription, ICESArea, Species, speciesCode)

# Check correct areas
wgbfas_tab <- dplyr::distinct(wgbfas_ok, SpeciesName, StockCode, ICESArea, StockKey)

wgbfas_tab_t <- tidyr::spread(wgbfas_tab, key = ICESArea, value = StockKey)

rm(wgbfas_tab, wg_dc, wg_eg, wg, wgbfas_tab_t)
```


## WGNSSK
```{r}
wg_name <- "WGNSSK"
wg <- subset(stocks, EG == wg_name)
wg_dc <- subset(stocks_data_call, `Expert Group` == wg_name)
wg_eg <- subset(stocks_sid, ExpertGroup == wg_name)

# Check from stock with overlapping areas, species
dup <- mutate(group_by(wg, SpeciesName, ICESArea), no = length(StockCode))
nrow(subset(dup, no > 1))

rm(dup)

## Remove other problems
wg <- subset(wg, !(StockCode == "ple.27.21-32" & ICESArea == "27.3.a.20"))

# Check that all data call and sid stocks are included
mis_eg <- subset(wg_eg, !(StockKeyLabel %in% wg$StockCode))
unique(mis_eg$StockKeyLabel)
mis_dc <- subset(wg_dc, !(`Stock Key Label` %in% wg$StockCode))
unique(mis_dc$`Stock Key Label`)

rm(mis_eg, mis_dc)

wgnssk_ok <- distinct(wg, SpeciesName, StockCode, EG, StockKey, 
StockKeyDescription, ICESArea, Species, speciesCode)

# Check correct areas
wgnssk_tab <- dplyr::distinct(wgnssk_ok, SpeciesName, StockCode, ICESArea, StockKey)

wgnssk_tab_t <- tidyr::spread(wgnssk_tab, key = ICESArea, value = StockKey)

rm(wgnssk_tab, wgnssk_tab_t, wg_dc, wg_eg, wg)
```

## NWWG
```{r}
wg_name <- "NWWG"
wg <- subset(stocks, EG == wg_name)
wg_dc <- subset(stocks_data_call, `Expert Group` == wg_name)
wg_eg <- subset(stocks_sid, ExpertGroup == wg_name)

# Check from stock with overlapping areas, species
dup <- mutate(group_by(wg, SpeciesName, ICESArea), no = length(StockCode))
nrow(subset(dup, no > 1))

wg <- subset(wg, !(StockCode == "cod.27.24-32" & ICESArea == "27.3.d.24"))
wg <- subset(wg, !(Species == "QDQ"))

rm(dup)

## Remove other problems
wg <- subset(wg, !(StockCode == "ple.27.21-32" & ICESArea == "27.3.a.20"))

# Check that all data call and sid stocks are included
mis_eg <- subset(wg_eg, !(StockKeyLabel %in% wg$StockCode))
unique(mis_eg$StockKeyLabel)
mis_dc <- subset(wg_dc, !(`Stock Key Label` %in% wg$StockCode))
unique(mis_dc$`Stock Key Label`)

rm(mis_eg, mis_dc)

nwwg_ok <- distinct(wg, SpeciesName, StockCode, EG, StockKey, 
StockKeyDescription, ICESArea, Species, speciesCode)

# Check correct areas
wgnssk_tab <- dplyr::distinct(nwwg_ok, SpeciesName, StockCode, ICESArea, StockKey)

wgnssk_tab_t <- tidyr::spread(wgnssk_tab, key = ICESArea, value = StockKey)

rm(wgnssk_tab, wgnssk_tab_t, wg_dc, wg_eg, wg)
```


## WGWIDE
```{r}
wg_name <- "WGWIDE"
wg <- subset(stocks, EG == wg_name)
wg_dc <- subset(stocks_data_call, `Expert Group` == wg_name)
wg_eg <- subset(stocks_sid, ExpertGroup == wg_name)

# Check from stock with overlapping areas, species
dup <- mutate(group_by(wg, SpeciesName, ICESArea), no = length(StockCode))
nrow(subset(dup, no > 1))

wg <- subset(wg, !(StockCode == "cod.27.24-32" & ICESArea == "27.3.d.24"))
wg <- subset(wg, !(Species == "QDQ"))

rm(dup)

## Remove other problems
wg <- subset(wg, !(StockCode == "ple.27.21-32" & ICESArea == "27.3.a.20"))

# Check that all data call and sid stocks are included
mis_eg <- subset(wg_eg, !(StockKeyLabel %in% wg$StockCode))
unique(mis_eg$StockKeyLabel)
mis_dc <- subset(wg_dc, !(`Stock Key Label` %in% wg$StockCode))
unique(mis_dc$`Stock Key Label`)

rm(mis_eg, mis_dc)

wgwide_ok <- distinct(wg, SpeciesName, StockCode, EG, StockKey, 
StockKeyDescription, ICESArea, Species, speciesCode)

# Check correct areas
wgnssk_tab <- dplyr::distinct(wgwide_ok, SpeciesName, StockCode, ICESArea, StockKey)

wgnssk_tab_t <- tidyr::spread(wgnssk_tab, key = ICESArea, value = StockKey)

rm(wgnssk_tab, wgnssk_tab_t, wg_dc, wg_eg, wg)
```

# Combine data
```{r}

relation <- rbind(hawg_ok, wgbfas_ok, wgnssk_ok)

```


# Check areas
## Add overlying
```{r}
ices_area_codes <- unique(ices_area$Key)

relation_area_over <- relation
added_areas <- c()

for (i in c(1:length(ices_area_codes))) {
  
  area <- paste0(ices_area_codes[i], "\\.")
  areas_over <- stringr::str_subset(ices_area_codes, paste0("^", area))
  areas_over <- areas_over[areas_over != area]
  
  area_relation <- subset(relation, ICESArea %in% areas_over)
  area_relation <- mutate(group_by(area_relation, StockCode), no_area = length(ICESArea))
  area_relation <- subset(area_relation, no_area == length(areas_over))
  area_relation$ICESArea <- gsub(pattern = "\\\\.", "", area)
  
  relation_area_over <- rbind(relation_area_over, select(area_relation, -no_area))
  
  added_areas <- rbind(added_areas, area_relation)

}

relation_area_over <- distinct(relation_area_over, SpeciesName, StockCode, EG, StockKey, 
StockKeyDescription, ICESArea, Species, speciesCode)

relation_area_over_tab <- dplyr::distinct(relation_area_over, SpeciesName, StockCode, ICESArea, StockKey)

relation_area_over_t <- tidyr::spread(relation_area_over_tab, key = ICESArea, value = StockKey)
```


## Add underlying
```{r}
ices_area_codes <- unique(ices_area$Key)

relation_area_under <- relation
added_areas <- c()

for (i in c(1:length(ices_area_codes))) {
  area <- ices_area_codes[i]
  areas_under <- stringr::str_subset(ices_area_codes, area)
  areas_under <- areas_under[areas_under != area]
  
  if(length(areas_under > 0)) {
  area_relation <- subset(relation_area_over, ICESArea %in% area)
  if (nrow(area_relation) > 0) {
    area_relation$ICESArea <- paste(areas_under, collapse = ",")
    area_relation <- tidyr::separate_longer_delim(data = area_relation, cols = ICESArea, ",")
    
    relation_area_under <- rbind(relation_area_under, area_relation)
    added_areas <- rbind(added_areas, area_relation)
  }
}}

relation <- rbind(relation_area_over, relation_area_under)

relation <- distinct(relation, SpeciesName, StockCode, EG, StockKey, 
StockKeyDescription, ICESArea, Species, speciesCode)

relation_area_under_tab <- dplyr::distinct(relation, SpeciesName, StockCode, ICESArea, StockKey)

relation_area_under_t <- tidyr::spread(relation_area_under_tab, key = ICESArea, value = StockKey)
```

# Add FMU to relation
```{r}
relation$FMU <- NA

relation$FMU[substr(relation$StockCode, 1, 6) == "san.sa"] <- 
  substr(relation$StockCode, 5, 10)[substr(relation$StockCode, 1, 6) == "san.sa"]

relation$FMU[relation$StockCode == "her.27.irls" & relation$ICESArea == "27.7.a"] <- 
  "27.7.a.s"

relation$FMU[relation$StockCode == "her.27.nirs" & relation$ICESArea == "27.7.a"] <- 
  "27.7.a.n"

relation$FMU[relation$StockCode == "her.27.6aS7bc" & relation$ICESArea == "27.6.a"] <- 
  "27.6.a.s"

relation$FMU[relation$StockCode == "her.27.6aN" & relation$ICESArea == "27.6.a"] <- 
  "27.6.a.n"

# relation$FMU[relation$StockCode == "her.27.1-24a514a" & relation$ICESArea == "27.4.a"] <- 
  # "27.6.a.n"


#her.27.1-24a514a


```


# Check relation per species
```{r, results='asis'}

dup <- mutate(group_by(relation, SpeciesName, ICESArea, FMU), no = length(StockCode))
nrow(subset(dup, no > 1))


```


# Export

```{r, eval = T}
dtu_aqua_stock_relation <- relation

usethis::use_data(dtu_aqua_stock_relation, internal = F, overwrite = T)


```
